### 2021-08-04 paper/paper.tex
remove colliding flows setup since it does not look that impressive for a general audience

We are broadly interested in simulation-focused scientific computing, in particular
in hyperbolic partial differential equations (PDEs), computational fluid dynamics (CFD),
and related problems. The focus of our research ranges from specific applications
in CFD to general multi-physics coupling strategies to the high-order numerical methods 
such simulations are built upon. In addition, we are academics and involved in teaching 
students in these areas of science. Thus, we would like to have a 
code that is
\begin{enumerate}
  \item extendable to be useful for research and development
  \item easy to apply/understand to be useful for students/collaborators
  \item fast enough to be useful beyond small toy problems
\end{enumerate}
Additionally, we are greedy\footnote{See \url{https://julialang.org/blog/2012/02/why-we-created-julia/}} 
and wish to include all these features within a single code.
Roughly one year of collaborative work has resulted in the current version of
\trixi\footnote{\url{https://github.com/trixi-framework/Trixi.jl}}, providing adaptive
high-order numerical simulations of hyperbolic partial differential equations (PDEs)
in Julia \cite{bezanson2017julia}. Starting as an experiment, we have been able
to reach more and more of our goals with \trixi.

@@

  \item High-order matrix-free discontinuous Galerkin methods
  \begin{itemize}
    \item Kinetic energy preserving and entropy-stable methods
    \item Entropy-stable shock-capturing
    \item Positivity-preserving limiting
  \end{itemize}

  \item Multiple governing equations
  \begin{itemize}
    \item Compressible Euler equations (optionally with self-gravity)
    \item Magnetohydrodynamics equations
    \item Acoustic perturbation equations
    \item Hyperbolic diffusion for elliptic problems
    \item Lattice-Boltzmann equations (D2Q9 and D3Q27 schemes)
    \item Several scalar conservation laws (advection, Burgers)
  \end{itemize}

@@

\trixi is based on the method of lines. Thus, a discretization of \eqref{eq:hcl}
is obtained in two steps. First, a spatial semidiscretization is created. Next,
the resulting ordinary differential equation (ODE) is solved using a time
integration method. Currently, \trixi focuses on the spatial semidiscretization
and uses mostly Runge-Kutta methods implemented in OrdinaryDiffEq.jl, a part
of DifferentialEquations.jl \cite{rackauckas2017differentialequations}.

\begin{figure*}[htbp]
  \includegraphics[width=\linewidth]{../figures/trixi_global_overview}
  \caption{Schematic overview of the basic components of \trixi and how they
           interact with each other.}
  \label{fig:trixi_global_overview}
\end{figure*}

Figure~\ref{fig:trixi_global_overview} presents an overview of the basic
components of \trixi. The most important structure is the semidiscretization,
which bundles all information about the spatial part. The mathematical-physical
model is determined by the \lstinline{equations}, the \lstinline{initial_condition},
\lstinline{boundary_conditions}, and possible \lstinline{source_terms}. The
\lstinline{solver} describes purely numerical parameters determining the specific
discretization method such as discontinuous Galerkin or finite difference
methods, kinetic energy preserving or shock capturing approaches. Finally, the
\lstinline{mesh} has a necessarily hybrid role including information about the
spatial domain and its discretization.

@@

The design of \trixi was guided by extensibility, ease of use, and efficiency
(in this particular order). Thus, it is likely possible to make (some parts) even 
faster at the cost of simplicity and readibility. Nevertheless, \trixi is already quite
fast, even compared to established high-performance open source software,
as described in details in Section~\ref{sec:performance-comparison}. In this
section, we focus mostly on the first two goals.

@@

The most important design decision was to create \trixi as a library. In the
context of simulation-focused scientific computing, there are a lot of (open
or closed source) codes that are designed as monolithic applications that can
simulate a specific setup. These codes can often be configured to a limited extent 
by specifying compile time options or editing parameter files. In contrast, users of 
\trixi do not need to compile anything or learn parameter file syntax. Instead, they can
just load the library and write their simulation setups in Julia. Scripts which describe
a simulation setup for \trixi are referred to as ``elixirs''. As of version v0.3.50, \trixi
comes with \ca 200 example elixirs. A basic elixir might look as follows:
% TODO: Update Trixi version above

@@

In traditional monolithic code bases, it is often necessary to implement a new
feature for all combinations of possible (compile time) options. This effort
can make it hard to experiment with new ideas. In contrast, Julia's dynamic
nature, multiple dispatch, and just-ahead-of-time compilation allow us to implement only the features
that are strictly necessary. For example, if a user wants to simulate a new
physics model only on Cartesian grids with smooth solutions, there is no need
to implement anything for curvilinear coordinates or shock capturing approaches.
In addition, it allows to extend \trixi step by step with new capabilities.
For example, there are ongoing efforts to incorporate summation by parts finite
difference methods via SummationByPartsOperators.jl \cite{ranocha2021sbp} and
discontinuous Galerkin methods on simplices via StartUpDG.jl\footnote{\url{https://github.com/jlchan/StartUpDG.jl}} 
in \trixi. This becomes feasible since not every feature combination needs to be considered, 
enabling developers to introduce variants of existing solvers on only a subset of the available
mesh types via multiple dispatch.

@@

Many scientists implementing numerical methods often have no real experience or
education in software development. Julia and its ecosystem support these
researches by making it easy to set up unit and regression tests, since a
testing framework is included in the standard library and default package
generation setups prepare the file structure accordingly. This allows test
driven development and continuous integration (CI), which makes code restructuring
and optimization feasible.

Julia itself and most packages are developed on GitHub. The community provides
a lot of tools for using corresponding CI setups, preparing and publishing
documentation\footnote{\url{https://github.com/JuliaDocs/Documenter.jl}},
and related tasks. This setup also encourages other good software development
practices that we use in \trixi, such as mandatory code review before pull requests
are merged.

@@

\subsubsection{Packages can really be used together at low (or no) cost}

Due to Julia's package manager, the cost of using external dependencies is low
enough to not be a problem. Traditionally, many scientific simulation codes tend
to reduce the number of external dependencies as much as possible due to the
complexity of handling different versions and making everything work together.
Such an approach typically leads to significant code duplication. For example, many
CFD codes implement their own time integration methods, arguing that the spatial
part is much more complex. While this is often true, it is still more efficient
in terms of human workload to be able to reuse existing implementations. This
allows expert on time integration methods to develop specialized algorithms and
implement them in open source software while practitioners or researchers focusing
on spatial semidiscretizations ca benefit immediately by changing only one line
of code. For example, optimized time integration methods were developed in
\cite{ranocha2021optimized} and implemented in OrdinaryDiffEq.jl. These methods
can be used with \trixi by changing a single line of code. In contrast, researchers
working without external dependencies would need to spend a day or two to digest
details before even beginning implementation. This additional overhead in developer
time makes it less likely that researchers would implement these methods in their own 
codes and benefit from recent algorithmic developments.

@@

Such a difference between the programming languages used in the front- and
back-ends lead to a well-known barrier between user and developers of a library.
Julia was designed from the beginning to make it possible to write both simple
high-level code and highly efficient compute kernels, thereby solving the
two-language problem --- at least to a significant extent. Highly efficient low-level
code will of course still look different from simple high-level code, but the
barrier between them is much smaller. In particular, this makes it easier for
users to transition to developers and contribute to a package. This also contributes
to the following observation:

@@

We have presented \trixi, a Julia library for adaptive high-order numerical
simulations of hyperbolic PDEs. As researchers in numerical analysis and
scientific computing, our goals were to create a framework that is extendable,
easy to use, and fast (in this particular order). Making use of Julia's strengths,
we have been successful based on recent publications making use of \trixi,
the number of students and researchers working with \trixi, and serial performance
comparisons with established Fortran codes. Having worked on \trixi for a bit
more than one year allows us to give an assessment of Julia for simulation-focused
scientific computing.

>>>>

We are broadly interested in simulation-focused scientific computing, in particular
numerical approximations for hyperbolic partial differential equations (PDEs), computational fluid dynamics (CFD),
and related problems. The focus of our research ranges from specific applications
in CFD to general multi-physics coupling strategies to the high-order numerical methods 
such simulations are built upon. In addition, we are academics and involved in teaching 
students in these areas of science. Thus, we would like to have a 
code that is
\begin{enumerate}
  \item extendable to be useful for research and development
  \item easy to apply/understand to be useful for students/collaborators
  \item fast enough to be useful beyond small toy problems
\end{enumerate}
Additionally, we are greedy\footnote{See \url{https://julialang.org/blog/2012/02/why-we-created-julia/}} 
and wish to include all these features within a single code.
Roughly one year of collaborative work has resulted in the current version of
\trixi\footnote{\url{https://github.com/trixi-framework/Trixi.jl}}, providing adaptive
high-order numerical simulations of hyperbolic partial differential equations (PDEs)
in Julia \cite{bezanson2017julia}. Starting as an experiment, we have been able
to reach more and more of our goals with \trixi.

@@

  \item High-order matrix-free discontinuous Galerkin (DG) methods
  \begin{itemize}
    \item Kinetic energy preserving and entropy-stable methods
    \item Entropy-stable shock-capturing
    \item Positivity-preserving limiting
  \end{itemize}

  \item Multiple governing equations
  \begin{itemize}
    \item Compressible Euler equations (optionally with self-gravity)
    \item Magnetohydrodynamics (MHD) equations
    \item Multicomponent compressible Euler and MHD equations
    \item Acoustic perturbation equations
    \item Hyperbolic diffusion for elliptic problems
    \item Lattice-Boltzmann equations (D2Q9 and D3Q27 schemes)
    \item Several scalar conservation laws (\eg, advection, Burgers)
  \end{itemize}

@@

\trixi is built from the method of lines. Thus, a discretization of \eqref{eq:hcl}
is obtained in two steps. First, a spatial semidiscretization is created. Next,
the resulting ordinary differential equation (ODE) is solved using a time
integration method. Currently, \trixi focuses on the spatial semidiscretization
and uses mostly Runge-Kutta methods implemented in OrdinaryDiffEq.jl, a part
of DifferentialEquations.jl \cite{rackauckas2017differentialequations}.

\begin{figure*}[htbp]
  \includegraphics[width=\linewidth]{../figures/trixi_global_overview}
  \caption{Schematic overview of the basic components in \trixi and how they
           interact.}
  \label{fig:trixi_global_overview}
\end{figure*}

Figure~\ref{fig:trixi_global_overview} presents an overview of the basic
components of \trixi. The most important structure is the semidiscretization,
which bundles all information about the spatial approximation. The mathematical-physical
model is determined by the \lstinline{equations}, the \lstinline{initial_condition},
\lstinline{boundary_conditions}, and possible \lstinline{source_terms}. The
\lstinline{solver} describes purely numerical parameters determining the specific
discretization method such as discontinuous Galerkin or finite difference
methods, kinetic energy preserving or shock capturing approaches. Finally, the
\lstinline{mesh} has a necessarily hybrid role including information about the
spatial domain and its discretization.

@@

The design of \trixi was guided by extensibility, ease of use, and efficiency
(in this particular order). Thus, it is likely possible to make (some parts) even 
faster at the cost of simplicity and readibility. Nevertheless, \trixi is already quite
fast, even compared to established high-performance open source software,
as described in detail in Section~\ref{sec:performance-comparison}. In this
section, we focus mostly on the first two goals.

@@

The most important design decision was to create \trixi as a library. In the
context of simulation-focused scientific computing, there are many (open
or closed source) codes that are designed as monolithic applications that can
simulate a specific setup. These codes can often be configured to a limited extent 
by specifying compile time options or editing parameter files. In contrast, users of 
\trixi do not need to compile anything or learn parameter file syntax. Instead, they can
just load the library and write their simulation setups in Julia. Scripts which describe
a simulation setup for \trixi are referred to as ``elixirs''. As of version v0.3.50, \trixi
comes with \ca 200 example elixirs. A basic elixir might look as follows:
% TODO: Update Trixi version above

@@

In traditional monolithic code bases, it is often necessary to implement a new
feature for all combinations of possible (compile time) options. This effort
can make it difficult to experiment with new ideas. In contrast, Julia's dynamic
nature, multiple dispatch, and just-ahead-of-time compilation allow us to implement only the features
that are strictly necessary. For example, if a user wants to simulate a new
physics model only on Cartesian grids with smooth solutions, there is no need
to implement anything for curvilinear coordinates or shock capturing approaches.
In addition, it allows a user to extend \trixi step by step with new capabilities.
For example, there are ongoing efforts to incorporate summation by parts finite
difference methods via SummationByPartsOperators.jl \cite{ranocha2021sbp} and
discontinuous Galerkin methods on simplex elements via StartUpDG.jl\footnote{\url{https://github.com/jlchan/StartUpDG.jl}} in \trixi.
This is feasible because the modular design of \trixi gives users the flexibility to select
a subset of the available meshes/methods, selected via multiple dispatch,
to test their newly implemented feature(s).

@@

Many scientists implementing numerical methods often have no real experience or
education in software development. Julia and its ecosystem support these
researchers by making it easy to set up unit and regression tests, since a
testing framework is included in the standard library and default package
generation setups prepare the file structure accordingly. This allows test
driven development and continuous integration (CI), which makes code restructuring
and optimization straightforward.

Julia itself and most packages are developed on GitHub. The community provides
many tools to use corresponding CI setups to prepare and publish
documentation\footnote{\url{https://github.com/JuliaDocs/Documenter.jl}},
and related tasks. This setup also encourages other good software development
practices that we use in \trixi, such as mandatory code review before pull requests
are merged.

@@

\subsubsection{Packages can be used together at low (or no) cost}

Due to Julia's package manager, the cost of using external dependencies is low
enough to not be a problem. Traditionally, many scientific simulation codes tend
to reduce the number of external dependencies as much as possible due to the
complexity of handling different versions and making everything work together.
Such an approach typically leads to significant code duplication. For example, many
CFD codes implement their own time integration methods, arguing that the spatial
part is much more complex. While this is often true, it is still more efficient
in terms of human workload to be able to reuse existing implementations. This
allows experts on time integration methods to develop specialized algorithms and
implement them in open source software while practitioners or researchers focusing
on spatial semidiscretizations ca benefit immediately by changing only one line
of code. For example, optimized time integration methods were developed in
\cite{ranocha2021optimized} and implemented in OrdinaryDiffEq.jl. These methods
can be used with \trixi by changing a single line of code. In contrast, researchers
working without external dependencies would need to spend a day or two to digest
details before even beginning implementation. This additional overhead in developer
time makes it less likely that researchers would implement these methods in their own 
codes and benefit from recent algorithmic developments.

@@

Such a difference between the programming languages used in the front- and
back-ends lead to a well-known barrier between a user and the developers of a library.
Julia was designed from the beginning to make it possible to write both simple
high-level code and highly efficient compute kernels, thereby solving the
two-language problem --- at least to a significant extent. Highly efficient low-level
code will of course still look different from simple high-level code, but the
barrier between them is much smaller. In particular, this makes it easier for
users to transition to developers and contribute to a package. This also contributes
to the following observation:

@@

We have presented \trixi, a Julia library for adaptive high-order numerical
simulations of hyperbolic PDEs. As researchers in numerical analysis and
scientific computing, our goals were to create a framework that is extendable,
easy to use, and fast (in this particular order). Making use of Julia's strengths,
we have been successful based on recent publications making use of \trixi,
the number of students and researchers working with \trixi, and serial performance
comparisons with an established Fortran code. Having developed, from scratch, \trixi for a bit
more than one year allows us to give an assessment of Julia for simulation-focused
scientific computing.
